\section{Education}\label{education}

\begin{description}
\itemsep3pt\parskip0pt\parsep0pt
\item[2011--15]
Ph.D.\ Biology, Simon Fraser University, Burnaby, Canada\\ Variance and
extreme events in population ecology\\ Fulbright Scholar to the
University of Washington in 2012--13\\ Dean of Graduate Studies
Convocation Medal
\item[2008--10]
M.Sc.\ Biology, Dalhousie University, Halifax, Canada\\ Trends, drivers,
and ecosystem effects of expanding global invertebrate fisheries\\
Canadian \href{http://goo.gl/nA1zE}{Governor General's Academic Gold
Medal}
\item[2004--07]
B.Sc.\ (Hons.) Environmental Science, Dalhousie University, Halifax,
Canada\\ First Class Honours \& Environmental Programmes Honour Society
Medal
\item[2001--03]
B.Sc.\ Candidate (Hons.) Kinesiology, University of Waterloo, Waterloo,
Canada; transferred to Dalhousie University in 2004.
\end{description}

\section{Publications}\label{publications}

\begin{description}
\itemsep3pt\parskip0pt\parsep0pt
\item[2015]
Monnahan, C.C., K. Ono, S.C Anderson, M.B. Rudd, A.C. Hicks, F.
Hurtado-Ferro, K.F. Johnson, P.T. Kuriyama, R.R. Licandeo, C.C. Stawitz,
I.G. Taylor, J.L. Valero. The effect of length bin width on growth
estimation in integrated age-structured stock assessments. In press.
\url{http://doi.org/10.1016/j.fishres.2015.09.007}
\item[2015]
Kuriyamaa, P.T., F. Hurtado-Ferro, K. Ono, A.C. Hicks, I.G. Taylor, R.R.
Licandeo, K.F. Johnson, S.C. Anderson, C.C. Monnahan, M.B. Rudd, C.C.
Stawitz, J.L. Valero. An empirical weight-at-age approach reduces
estimation bias compared to modeling parametric growth in integrated,
statistical stock assessment models when growth is time varying. In
press. \url{http://doi.org/10.1016/j.fishres.2015.11.002}
\item[2015]
Finnegan*, S., \textbf{S.C. Anderson}*, P.G. Harnik*, C. Simpson, D.P.
Tittensor, J.E. Byrnes, Z.V. Finkel, D.R. Lindberg, L.H. Liow, R.
Lockwood, H.K. Lotze, C.M. McClain, J.L. McGuire, A. O'Dea, J.M.
Pandolfi. Paleontological baselines for evaluating extinction risk in
the modern oceans. Science. 348(6234): 567--570.
\url{http://doi.org/10.1126/science.aaa6635}
(\href{https://github.com/seananderson/paleobaselines}{GitHub
repository}) (\emph{*Authors contributed equally}).
\item[2015]
Orzechowski, E.A., R. Lockwood, J.E. Byrnes, \textbf{S.C. Anderson}, S.
Finnegan, Z.V. Finkel, P.G. Harnik, D.R. Lindberg, L.H. Liow, H.K.
Lotze, C.R. McClain, J.L. McGuire, A. O'Dea, J.M. Pandolfi, C. Simpson,
D.P. Tittensor. Marine extinction risk shaped by trait-environment
interactions over 500 million years. In press at Global Change Biology.
\item[2015]
Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down and bottom-up
forces interact at thermal range extremes on American lobster. Journal
of Animal Ecology. 83(3):840--850.
\url{http://doi.org/10.1111/1365-2656.12322}.
(\href{https://github.com/seananderson/lobsters-predators}{GitHub
repository})
\item[2015]
Hurtado-Ferro, F., C.S. Szuwalski, J.L. Valero, \textbf{S.C. Anderson},
C.J. Cunningham, K.F. Johnson, R.R. Licandeo, C.R. McGilliard, C.C.
Monnahan, M.L. Muradian, K. Ono, K.A. Vert-Pre, A.R. Whitten, A.E. Punt.
Looking in the rear-view mirror: bias and retrospective patterns in
integrated, age-structured stock assessment models. ICES Journal of
Marine Science. 72(1): 99--110.
\url{http://doi.org/10.1093/icesjms/fsu198}.
\item[2015]
\textbf{Anderson, S.C.}, J.W. Moore, M.M. McClure, N.K. Dulvy, A.B.
Cooper. Portfolio conservation of metapopulations under climate change.
Ecological Applications. 2015. 25(2): 559--572.
\url{http://doi.org/10.1890/14-0266.1}
(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2014_salmon_portfolios.pdf}{PDF},
\href{https://github.com/seananderson/metafolio}{GitHub repository},
\href{http://cran.r-project.org/web/packages/metafolio/index.html}{CRAN
R package}).
\item[2015]
Johnson, K.F., C.C. Monnahan, C.R. McGilliard, K.A. Vert-pre,
\textbf{S.C. Anderson}, C.J. Cunningham, F. Hurtado-Ferro, R.R.
Licandeo, M.L. Muradian, K. Ono, C.S. Szuwalski, J.L. Valero, A.R.
Whitten, A.E. Punt. Time-varying natural mortality in fisheries stock
assessment models: identifying a default approach. ICES Journal of
Marine Science. 72(1): 137--150.
\url{http://doi.org/10.1093/icesjms/fsu055}
(\href{http://icesjms.oxfordjournals.org/content/early/2014/04/09/icesjms.fsu055.full.pdf?keytype=ref\&ijkey=NEXmZIkz3289u3z}{PDF},
\href{https://github.com/ss3sim/natural-mortality}{GitHub repository}).
\item[2015]
Ono, K., R. Licandeo, M.L. Muradian, C.J. Cunningham, \textbf{S.C.
Anderson}, F. Hurtado-Ferro, K.F. Johnson, C.R. McGilliard, C.C.
Monnahan, C.S. Szuwalski, J.L. Valero, K.A. Vert-pre, A.R. Whitten, A.E.
Punt. The importance of length and age composition data in statistical
catch-at-age models for marine species. ICES Journal of Marine Science.
72(1): 31--43. \url{http://doi.org/10.1093/icesjms/fsu007}
(\href{https://dl.dropboxusercontent.com/u/254940/papers/Ono_etal_2014_importance_of_length_and_age_composition_data.pdf}{PDF}).
\item[2014]
Farmer, R.G., M.L. Leonard, J.E. Mills Flemming, \textbf{S.C. Anderson}.
Observer aging and long-term avian survey data quality. Ecology and
Evolution. 4(12): 2563--2576. \url{http://doi.org/10.1002/ece3.1101}
(\href{http://onlinelibrary.wiley.com/doi/10.1002/ece3.1101/pdf}{PDF}).
\href{http://news.nationalgeographic.com/news/2014/08/140805-aging-birders-breeding-bird-survey-volunteers-science/}{National
Geographic news story}.
\item[2014]
\textbf{Anderson, S.C.}, C.C. Monnahan, K.F. Johnson, K. Ono, J.L.
Valero. ss3sim: An R package for fisheries stock assessment simulation
with Stock Synthesis. PLOS ONE. 9(4): e92725.
\url{http://doi.org/10.1371/journal.pone.0092725}
(\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0092725\&representation=PDF}{PDF}).
(\href{http://cran.r-project.org/web/packages/ss3sim/index.html}{R
package}, \href{https://github.com/ss3sim/ss3sim}{code},
\href{https://dl.dropboxusercontent.com/u/254940/ss3sim-vignette.pdf}{vignette}).
\item[2014]
O'Regan, S.M., W.J. Palen, \textbf{S.C. Anderson}. Climate warming
mediates negative impacts of rapid pond drying for three amphibian
species. Ecology. 95(4): 845--855.
\url{http://doi.org/10.1890/13-0916.1}
(\href{http://www.esajournals.org/doi/pdf/10.1890/13-0916.1}{PDF}).
\href{http://www.esajournals.org/userimages/ContentEditor/1398099670397/i0012-9658-95-4-cov.pdf}{Cover
Article}.
\item[2013]
Favaro, B., D.C. Braun, \textbf{Earth2Ocean Research Derby}. Research
Derby: A pressure cooker for creative collaborative science. Ideas in
Ecology and Evolution. Ideas in Ecology and Evolution. 6: 40--46.
\url{http://doi.org/10.4033/iee.2013.6.9.n}
(\href{http://library.queensu.ca/ojs/index.php/IEE/article/download/4931/4899}{PDF}).
\item[2013]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. Ecological prophets:
Quantifying metapopulation portfolio effects. Methods in Ecology and
Evolution. 4(10): 971--981. \url{http://doi.org/10.1111/2041-210X.12093}
(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2013_ecological_prophets_with_SOM.pdf}{PDF},
\href{https://github.com/seananderson/ecofolio}{GitHub R package}).
\item[2013]
Artelle, K.A., \textbf{S.C. Anderson}, A.B. Cooper, P.C. Paquet, J.D.
Reynolds, C.T. Darimont. Confronting uncertainty in wildlife management:
performance of grizzly bear management. PLOS ONE. 8(11): e78041.
\url{http://doi.org/10.1371/journal.pone.0078041}
(\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0078041\&representation=PDF}{PDF}).
\item[2013]
Phillis*, C.C., S.M. O'Regan*, S.J. Green*, J.E. Bruce*, \textbf{S.C.
Anderson}, J.N. Linton, E.R. Derby, B. Favaro. Multiple pathways to
conservation success. Conservation Letters. 6(2): 98--106.
\url{http://doi.org/10.1111/j.1755-263X.2012.00294.x}
(\href{https://dl.dropboxusercontent.com/u/254940/papers/Phillis_etal_2012_Multiple_pathways_to_conservation_success.pdf}{PDF},
\href{https://github.com/seananderson/conservation_pathways}{GitHub
repository}). (*Authors contributed equally, listed in reverse
alphabetical order).
\item[2012]
Harnik, P.G., H.K. Lotze, \textbf{S.C. Anderson}, Z.V. Finkel, S.
Finnegan, D.R. Lindberg, L.H. Liow, R. Lockwood, C.R. McClain, J.L.
McGuire, A. O'Dea, J.M. Pandolfi, C. Simpson, D.P. Tittensor.
Extinctions in ancient and modern seas. Trends in Ecology and Evolution.
27(11): 608--617. \url{http://doi.org/10.1016/j.tree.2012.07.010}
(\href{https://dl.dropboxusercontent.com/u/254940/papers/Harnik_etal_2012_Extinctions_in_ancient_and_modern_seas.pdf}{PDF}).
\item[2012]
\textbf{Anderson, S.C.}, T.A. Branch, D. Ricard, H.K. Lotze. Assessing
global marine fishery status with a revised dynamic catch-based method
and stock-assessment reference points. ICES Journal of Marine Science.
69(8): 1491--1500. \url{http://doi.org/10.1093/icesjms/fss105}
(\href{http://icesjms.oxfordjournals.org/cgi/reprint/fss105?ijkey=pFRzMGEDycomI3X\&keytype=ref}{PDF}).
\item[2011]
\textbf{Anderson, S.C.}, R.G. Farmer, F. Ferretti, A.L.S. Houde, J.A.
Hutchings. Correlates of vertebrate extinction risk in Canada.
BioScience. 61(7): 538--549.
\url{http://doi.org/10.1525/bio.2011.61.7.8}
(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2011_BioScience_with_supplement.pdf}{PDF}).
\item[2011]
\textbf{Anderson, S.C.}, J.E. Mills Flemming, R. Watson, H.K. Lotze.
Serial exploitation of global sea cucumber fisheries. Fish and
Fisheries. 12(3): 317--339.
\url{http://doi.org/10.1111/j.1467-2979.2010.00397.x}
(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2011_seacucumbers_with_supplement.pdf}{PDF}).
\href{http://www.sciencemag.org/content/331/6014/129.1.full}{Featured in
Science}.
\item[2011]
Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down interactions
and temperature control of snow crab abundance in the northwest Atlantic
Ocean. Marine Ecology Progress Series. 429: 169--183.
\url{http://doi.org/10.3354/meps09081}
(\href{http://www.int-res.com/articles/meps_oa/m429p169.pdf}{PDF}).
\item[2011]
\textbf{Anderson, S.C.}, J.E. Mills Flemming, R. Watson, H.K. Lotze.
Rapid global expansion of invertebrate fisheries: trends, drivers, and
ecosystem effects. PLOS ONE. 6(3): e14735.
\url{http://doi.org/10.1371/journal.pone.0014735}
(\href{http://www.plosone.org/article/fetchObject.action?uri=info\%3Adoi\%2F10.1371\%2Fjournal.pone.0014735\&representation=PDF}{PDF}).
\item[2008]
\textbf{Anderson, S.C.}, H.K. Lotze, N.L. Shackell. Evaluating the
knowledge base for expanding low-trophic-level fisheries in Atlantic
Canada. Canadian Journal of Fisheries and Aquatic Sciences. 65(12):
2553--2571. \url{http://doi.org/10.1139/F08-156}
(\href{https://dl.dropboxusercontent.com/u/254940/papers/Anderson_etal_2008_knowledge.pdf}{PDF}).
Eighth most-read CJFAS article in 2008.
\end{description}

\section{Awards and scholarships}\label{awards-and-scholarships}

\begin{description}
\itemsep3pt\parskip0pt\parsep0pt
\item[2015]
David H. Smith Conservation Postdoctoral Research Fellowship
\item[2014]
Garfield Weston Foundation / BC Packers Ltd. Graduate Fellowship in
Marine Sciences
\item[2014]
Graduate Fellowship (two semesters), Simon Fraser University
\item[2012--13]
Canadian Fulbright Scholar award to the University of Washington in
2012--13
\item[2011]
Canadian \href{http://goo.gl/nA1zE}{Governor General's Academic Gold
Medal} for the top-ranked Master's Natural Sciences and Engineering
thesis at Dalhousie University in 2010
\item[2011--14]
Provost Prize of Distinction, Simon Fraser University
\item[2011--14]
Natural Sciences and Engineering Research Council of Canada Postgraduate
Scholarship (Doctoral)
\item[2007--10]
Faculty Research Grant Scholarship, Dalhousie University
\item[2007--09]
Graduate Studies Scholarship, Dalhousie University
\item[2007]
Environmental Programmes Honour Society Medal, Dalhousie University
\end{description}

\section{Software}\label{software}

\begin{description}
\itemsep3pt\parskip0pt\parsep0pt
\item[2014]
\textbf{Anderson, S.C.}, J.W. Moore, M.M McClure, N.K. Dulvy, A.B.
Cooper.\\ metafolio: Metapopulation simulations for conserving salmon
through portfolio optimization.
\url{http://cran.r-project.org/package=metafolio}
\item[2013]
\textbf{Anderson, S.C.}, C.C. Monnahan, K.F. Johnson, K. Ono, J.L.
Valero, C.J Cunningham, F. Hurtado-Ferro, R. Licandeo, C.R. McGilliard,
C.S. Szuwalski, K.A. Vert-pre, A.R. Whitten. ss3sim: Fisheries stock
assessment simulation testing with Stock Synthesis.
\url{http://cran.r-project.org/package=ss3sim}
\item[2013]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. ecofolio: Tools to
quantify metapopulation portfolio effects.
\url{https://github.com/seananderson/ecofolio}
\end{description}

\section{Invited talks and conference
presentations}\label{invited-talks-and-conference-presentations}

\begin{description}
\itemsep3pt\parskip0pt\parsep0pt
\item[2013]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. False prophets: The
challenges of quantifying ecological portfolios
(\href{http://seananderson.ca/talks/2013/PE_SAFS_quantsem.pdf}{slides}),
Quantitative Seminar, School of Aquatic and Fishery Sciences, University
of Washington, Seattle, WA, United States.
\item[2012]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. False prophets: The
ecological portfolio effect overestimates the benefit of diversity.
Branch Lab, School of Aquatic and Fishery Sciences, University of
Washington, Seattle, WA, United States.
\item[2012]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. Metapopulation
dynamics and the generality of the ecological portfolio effect. Les
Ecologistes Departmental Seminar, Simon Fraser University, Burnaby, BC,
Canada.
\item[2010]
\textbf{Anderson, S.C.} The rise of invertebrates, the fall of sea
cucumbers, and the risk of maturing late. Earth2Ocean Research Group,
Simon Fraser University, Burnaby, BC, Canada.
\item[2009]
\textbf{Anderson, S.C.}, J.E. Mills Flemming, R. Watson, H.K. Lotze.
Ecosystem impacts of the global expansion of invertebrate fisheries.
International Oceans Past II Conference \emph{Multidisciplinary
Perspectives on the History and Future of Marine Animal Populations}.
Vancouver, BC, Canada.
\item[2009]
\textbf{Anderson, S.C.}, J.E Mills Flemming, R. Watson, H.K. Lotze.
Global invertebrate fisheries: trends and consequences. NCEAS (National
Center for Ecological Analysis and Synthesis) Working Group
\emph{Finding Common Ground in Marine Conservation and Management}.
Santa Barbara, CA, USA.
\item[2009]
\textbf{Anderson, S.C.}, H.K. Lotze, N.L. Shackell. Evaluating the
knowledge base for expanding low-trophic level fisheries in Atlantic
Canada. Harvest Fisheries Seminar Series. Bedford Institute of
Oceanography, Fisheries and Oceans, Dartmouth, NS, Canada.
\item[2008]
\textbf{Anderson, S.C.}, H.K. Lotze, N.L. Shackell. Evaluating the
knowledge base for expanding low-trophic level fisheries in Atlantic
Canada. Departmental Seminar, Biology Department, Dalhousie University,
Canada.
\end{description}

\section{Teaching}\label{teaching}

\begin{description}
\itemsep3pt\parskip0pt\parsep0pt
\item[2014--15]
Developed self-directed lecture and exercises on ggplot2 for FISH 554:
Beautiful graphics in R, School of Aquatic and Fishery Sciences,
University of Washington, Seattle, WA, United States.\\
\url{http://seananderson.ca/ggplot2-FISH554/}
\item[2013--14]
Organizer of Stats Beerz --- a statistical help group attended by
graduate students and postdocs primarily in the Earth to Oceans research
group, but also the wider SFU Biology and Geography Departments, and the
School of Resource and Environmental Management (REM).
\item[2013]
Two-part workshop on data manipulation for Stats Beerz and Earth to
Oceans groups at Simon Fraser University with approximately 25
participants. An introduction to plyr, advanced concepts with plyr and
function debugging, and an introduction to dplyr.\\
\url{https://github.com/seananderson/plyr-statsbeerz}
\item[2013]
Instructor for BISC-888-1: Data Wrangling and Visualization in R, a
graduate-level course at Simon Fraser University, Burnaby, BC, Canada
with approximately 20 participants. Co-developed curriculum and
developed/delivered lectures, exercises, notes, and assignments for
three of six two-hour classes.\\
\url{https://github.com/seananderson/datawranglR} (see classes 03, 04,
05)
\item[2012]
Introduction to ggplot2.
(\href{http://seananderson.ca/courses/12-ggplot2/ggplot2_notes.pdf}{notes},
\href{http://seananderson.ca/courses/12-ggplot2/ggplot2_slides_with_examples.pdf}{slides})
Lecture for FISH 507H: Beautiful Graphics in R, School of Aquatic and
Fishery Sciences, University of Washington, Seattle, WA, United States
\item[2012]
Workshop on the R package plyr.
(\href{http://seananderson.ca/courses/12-plyr/plyr_2012.pdf}{notes},
\href{http://seananderson.ca/courses/12-plyr/plyr_2012_slides.pdf}{slides},
\href{http://seananderson.ca/courses/12-plyr/plyr_2012_examples.html}{examples})
Branch Lab, School of Aquatic and Fishery Sciences, University of
Washington, Seattle, WA, United States
\item[2011]
Multipanel plotting in R with base graphics.
(\href{http://seananderson.ca/courses/11-multipanel/multipanel.pdf}{notes},
\href{http://seananderson.ca/courses/11-multipanel/multipanel-slides.pdf}{slides}).
Earth2Ocean Research Group, Simon Fraser University, Burnaby, BC,
Canada.
\item[2011]
A brief introduction to R.
(\href{http://seananderson.ca/courses/11-rintro/RIntro.pdf}{notes},
\href{http://seananderson.ca/courses/11-rintro/RIntro.R}{workshop
code}). Earth2Ocean Research Group, Simon Fraser University, Burnaby,
BC, Canada.
\item[2011]
Workshop on the R package plyr
(\href{http://seananderson.ca/courses/11-plyr/plyr.pdf}{notes},
\href{http://seananderson.ca/courses/11-plyr/plyr-slides.pdf}{slides}).
Earth2Ocean Research Group, Simon Fraser University, Burnaby, BC,
Canada.
\item[2007--08]
Teaching Assistant, Organismal Biology and Ecology, Dalhousie
University, two semesters (BIOL 1021).
\item[2007--08]
Teaching Assistant, Marine Mammology, Dalhousie University (BIOL 4060).
\end{description}

\section{Contract employment}\label{contract-employment}

\begin{description}
\itemsep3pt\parskip0pt\parsep0pt
\item[2013]
Co-developed global fisheries status maps for the
\href{http://ramlegacy.org/}{RAM Legacy Stock Assessment Database} for
R. Hilborn
\item[2010]
Assessment of Arctic surfclam on Banquereau Bank using
industry-collected data for Clearwater Seafoods, Bedford, NS, Canada.
\item[2010]
Modelling length-frequency distribution changes of Atlantic halibut for
NAFO Divisions 4VWX5Z and 3NOP for Dr.\ K. Trzcinski, Population Ecology
Division, Bedford Institute of Oceanography, Fisheries and Oceans,
Dartmouth, NS, Canada.
\end{description}

\section{Working groups and
workshops}\label{working-groups-and-workshops}

\begin{description}
\itemsep3pt\parskip0pt\parsep0pt
\item[2011--13]
NESCent (National Evolutionary Synthesis Center, Durham, NC) Working
Group \emph{Determinants of Extinction in Ancient and Modern Seas} led
by Paul Harnik, Seth Finnegan, and Rowan Lockwood.
(\href{http://www.nescent.org/science/awards_summary.php?id=256}{URL})
\item[2010]
Atlantic Halibut Assessment Science Peer Review Meeting, Fisheries and
Oceans, Dartmouth, NS, Canada.
\item[2007--09]
NCEAS (National Center for Ecological Analysis and Synthesis, Santa
Barbara, CA) Distributed Graduate Seminar, in association with the
Working Group \emph{Finding Common Ground in Marine Conservation and
Management} led by Ray Hilborn and Boris Worm.
(\href{http://www.nceas.ucsb.edu/projects/12307}{URL})
\item[2008]
Workshop on Canadian Science and Management Strategies for Sea Cucumber
(\emph{Cucumaria frondosa}), Fisheries and Oceans, Dartmouth, NS, Canada
(work presented \emph{in absentia}).
\item[2007]
Workshop on Canadian Science and Management Strategies for Atlantic
Hagfish, Fisheries and Oceans, Dartmouth, NS, Canada.
(\href{http://www.dfo-mpo.gc.ca/CSAS/Csas/Publications/Pro-CR/2009/2009_009_e.htm}{URL})
\end{description}

\section{Reviews}\label{reviews}

Reviewer for Science, Ecology, Conservation Biology, Fish and Fisheries,
International Journal of Tropical Biology and Conservation, Journal of
Environmental Management, Endangered Species Research, Aquatic
Conservation
