# plyr examples

Let's load plyr and the data:

```{r}
library(plyr)
load("imdb.rda")
```

Let's start with a basic example. How does the mean movie length change through time?

```{r}
x0 <- ddply(imdb, "year", summarize, mean.length = mean(length, na.rm = TRUE))
plot(x0)
```

We can also look at the mean length by genre and year. We can work group by multiple columns using the c("column1", "column2") format.

```{r}
 x1 <- ddply(imdb, c("genre", "year"), summarize, mean.length = mean(length, na.rm = TRUE))

par(mfrow = c(5,5), mar = c(0,1.2,0,1), oma = c(3,3,1,1), cex = 0.6)
d_ply(x1, "genre", transform, {
  plot(year, mean.length, type = "l", xaxt = "n", xlim = c(1920, 2011))
  mtext(unique(genre), line = -1.5)
})
```

What if we wanted to detrend each time series with a linear model?

```{r}
x2 <- ddply(subset(x1, !is.na(mean.length)), "genre", transform, residual = residuals(lm(mean.length~year)))
head(x2)
```

How about running a regression of movie length and rating for each year and storing the models?

```{r}
x3 <- dlply(subset(imdb, year > 1950 & genre != "Short"), "year", function(x) 
  lm(length~imdb_rating, data = x))
x3[[1]]
```

Now what if we wanted to extract the slopes and their confidence intervals from those models and return them in a data frame? You might want to step into the function with browser() to help write the code.

```{r}
x4 <- ldply(x3, function(x) {
      est <- coef(x)[2]
      ci <- confint(x)
      ci.l <- ci[2, 1]
      ci.u <- ci[2, 2]
      data.frame(est, ci.l, ci.u)
})
library(ggplot2)
ggplot(subset(imdb, year >= 1990 & genre != "Short"), aes(imdb_rating, length)) + geom_point(alpha = 0.05) + facet_wrap(~year) + stat_smooth(method = "lm")
par(mfrow = c(1, 1))
with(x4, plot(year, est, ylim = c(min(ci.l), max(ci.u))))
d_ply(x4, "year", transform, segments(year, ci.l, year, ci.u))
```

